---
title: "individual1"
author: "Addison Schmidt"
date: "10/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
defense = read_excel('./cyclonesFootball2020.xlsx', sheet='Defensive')
str(defense)
offense = read_excel('./cyclonesFootball2020.xlsx', sheet='Offensive')
str(offense)
biography = read_excel('./cyclonesFootball2020.xlsx', sheet='Biography')
str(biography)
```
**Part One**
1.
```{r}
defense$Name = factor(defense$Name)
defense$Opponent_Opponent = factor(defense$Opponent_Opponent)
offense$Name = factor(offense$Name)
offense$Opponent_Opponent = factor(offense$Opponent_Opponent)
biography$Name = factor(biography$Name)
offense
defense
biography
```
2.
```{r}
offense = offense %>% 
  mutate(across(Receiving_REC:Passing_INT, as.numeric))
defense = defense %>% 
  mutate(across(Tackles_Solo:Pass_PB, as.numeric))
biography$Weight = as.numeric(biography$Weight)
```

3.
```{r}
biography = biography %>%
  separate(Height, c('ft', 'in'))
biography$ft = as.numeric(biography$ft)
biography$`in` = as.numeric(biography$`in`)
biography = biography %>%
  mutate(Height = ft*12 + `in`)
biography
```

4.
```{r}
offClean = offense %>%
  group_by(Name, Opponent_Opponent) %>%
  mutate(gameNum = row_number())
defClean = defense %>%
  group_by(Name, Opponent_Opponent) %>%
  mutate(gameNum = row_number())
bioClean = biography
str(offClean)
str(defClean)
str(bioClean)
```

**Part Two**
1.
```{r}
offClean = offClean %>%
  pivot_longer(Receiving_REC:Passing_INT, names_to = 'stat', values_to = 'value')
offClean = na.omit(offClean)
offClean
```
2.
```{r}
offStats = offClean %>%
  group_by(Name, stat) %>%
  summarise(sum = sum(value))
offStats
```

3. Noticeably, both the range and interquartile range of Passing_YDS dwarfs that of all the other stats. Receiving_YDS has the most outliers, with two.
```{r}
ggplot(offStats, aes(x=sum)) + geom_boxplot() + facet_wrap(~stat, scales="free")
```

4. The players scored better against Oklahoma, with the highest scoring player getting nearly 600 against Oklahoma, with just over 150 against Oregon.
```{r}
offStatsOpp = offClean %>%
  group_by(Name, stat, Opponent_Opponent) %>%
  summarise(sum = sum(value))
offStatsOppWide = offStatsOpp %>%
  pivot_wider(names_from=Opponent_Opponent, values_from=sum)
ggplot(offStatsOppWide, aes(x=Oregon, y=Oklahoma)) + geom_point()
```

4. (the second one)
```{r}
bioClean = bioClean %>%
  separate(Hometown, c('city', 'state'), sep=', ')
```

5.
```{r}
bioClean %>%
  group_by(state) %>%
  summarise(count = n()) %>%
  arrange(-count)
```

6. Brock Purdy was at the top range of players for all applicable stats, except for one outlier in rushing stats who was outside of the normal range.
```{r}
offStats
brockPurdy = offStats %>%
  filter(Name == "Purdy, Brock")
brockPurdy
  
ggplot(offStats, aes(x=sum)) + geom_boxplot() + geom_boxplot(data = brockPurdy, col="red") + facet_wrap(~stat, scales="free")
```